\documentclass[12pt, a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage[left=4cm, right=4cm]{geometry}
\usepackage{setspace}

\renewcommand{\labelitemi}{-}

\newenvironment{itemlist}
{
  \vspace{-0.5\topsep}
  \begin{itemize}
    \setlength{\itemsep}{4pt}
    % \setlength{\topsep}{0pt}
    \setlength{\parskip}{0pt}
    % \setlength{\partopsep}{0pt}
} {
  \end{itemize}
  \vspace{-0.5\topsep}
}

\onehalfspacing

\title{Architecture and Development Documentation of Nebula}
\author{Piotr Kocia}
\date{}

\begin{document}
\maketitle

\tableofcontents

\section{Introduction}
The purpose of this document is to explain the architecture and architectural
decisions, the development process and the structure of Nebula. The aim is to
shed light on what Nebula will be at the end of the planned development, what
features will be present and what risks are possible throughout the entire
process. Additionally, this document steps into some lower level architectural
details of certain systems so as to elucidate why certain decisions have been
made.

\section{Nebula}
Nebula is a standalone desktop program for designing, building and simulating
integrated electronic circuits (circuits). The end goal is to be able to create
circuits from basic logic building blocks (logic gates, gates) by connecting
them with wires, simulate their working by providing binary inputs and transform
the representation into technical diagrams within the program itself. The
program will be run as a standalone desktop window program where the user
interacts with it via user interface elements. The core planned features are:
\begin{itemlist}
\item realtime rendering of the circuits,
\item realtime interaction and modification of the circuits,
\item grouping of components into larger units for reuse,
\item simulation of execution for given binary inputs,
\item clock-based continuous simulation,
\item persistence of created designs.
\end{itemlist}

Several interesting features are sadly out of the scope of this project at the
time due to the imposed timeframe, hence we consider them extension features
which will be implemented once all core features are completed and time budget
allows. Such features are:
\begin{itemlist}
\item automatic layouting,
\item generation of technical diagrams.
\end{itemlist}

\section{Development Phases}
The development is divided into 4 phases: setup, bootstrap, core development,
finalisation.

\subsection{Setup}
In the setup phase the goal is to create a formalise the processes within the
project. This encompases initial project setup, standardisation and
establishment of guidelines and preparation of workflows, as well as environment
setup, acquisition of tools and brief technical onboarding.

A hard time estimate for this phase is 1 week with no planned overruns.

\subsection{Bootstrap}
In the bootstrap phase the goal is to create a working foundation of the project
which will enable further concurrent development of the planned features. This
phase consists primarily of low-level design of the critical systems and
modules, as well as early development of the blocking parts of the program.

A hard time estimate for this phase is 1 week with no planned overruns.

\subsection{Core Development}
Core development is the largest phase of the project optimistically estimated to
last 9 weeks. After that time the project enters the finalisation phase
regardless of whether an overrun occurs.

The main focus is on developing the core set of features of the application as
those have the highest priority and are essential to proper functioning and
experience of Nebula. In the unlikely case that all core features are
completed before the end of this phase, work on extension features may start.

\subsection{Finalisation}
The finalisation phase is the last phase of the project. This phase serves
primarily as a buffer for an overrun in the core development phase so that all
core features might be completed and polished. Any extension features
successfully developed before the end of this phase will be merged. The planned
duration of this phase is 1 week with a possible extension to 2 weeks.

\section{Architecture Overview}
Nebula consists of several interoperating modules: model, renderer, executor,
persistence, user interface.

DIAGRAM HERE

\subsection{Model}
Model is the core module responsible for in-memory storage of the logical and
structural representation of the circuits and the related functionality. The
Core Representation (CR) encompasses gates, components, IO connectors, wires.

Gates are builtin atomic components which perform boolean operations on inputs.
The only necessary gate is NAND, however, for the convenience of a user and
performance all basic operations are included (NOT, AND, NAND, OR, NOR, XOR,
XNOR). Additionally, SR Latch and D Latch may be considered for inclusion in
this set, as those are elemental memory units.

Components are groups of smaller components with defined IO acting as a single
building block.

IO connectors are used to transfer into and out of components. Types of IO
connectors are essential to proper UX as errors resulting from mismatched types
of IO connectors, such as input-input connections or 8 bit to 1 bit connection,
will be difficult to locate by the end user. No debugging functionality is
planned, hence minimising possible points of failure is crucial.

Wires are an explicit representation of the connections between components.
Wires may have junctions that split the wire into multiple lines transferring
the same value. There is no distinction between different bit width buses to
simplify construction of more complex circuits.

The CR, for the purposes of execution, may be lowered to a performance oriented
representation (PR) which is built on-demand before execution and is
structurally immutable.

\subsection{Renderer}
The renderer is fed geometry and color/texture data, and draws it on screen. The
internal design is entirely dependent on the graphics API, however, the
interface of this module should be straightforward in the way that it should be
possible to issue individual draws efficiently. It is an independent module that
does not have any strict ties to any other modules.

\subsection{Windowing}
Windowing is responsible for managing the application window as well as serving
as the first layer of user input handling. Primary responsibilities:
\begin{itemlist}
  \item processing user input into a manageable format,
  \item presenting the main window,
  \item handling interactions with the main window,
  \item providing a rendering context.
\end{itemlist}

\subsection{Executor}
Executor simulates the circuit with a set of given inputs propagating the
results of calculations throughout the system. Primary actions are:
\begin{itemlist}
  \item continuous clock synchronised execution at a given frequency,
  \item single-step clock synchronised execution,
  \item one-time execution on inputs change.
\end{itemlist}

\subsection{Persistence}
Storage of created designs is essential, as recreating them from scratch each
time would be deterring to say the least. Persistence must be capable of
serialising the model to store it externally and reconstruct the model from
serialised data. The primary actions are:
\begin{itemlist}
  \item export a component,
  \item import a component,
  \item save a project,
  \item load a project.
\end{itemlist}

\subsection{User Interface}
User Interaface (UI) is the primary method of interaction with the program. UI
exposes interactivity points that manipulate the state of the program and
handles user input, possible actions being:
\begin{itemlist}
  \item panning view,
  \item adding a component
  \item moving a component,
  \item creating a wire connection between two IO connectors,
  \item adding a wire junction,
  \item box-selecting gates,
  \item grouping gates into a larger component,
  \item starting simulation,
  \item importing/exporting components.
\end{itemlist}
UI should expose 2 views - the main view of the circuit and the component editor
view. The component editor is used to edit components, that is modify their
internal structure and modify IO. In the case that a component has its IO
edited, an old version of the component should be maintained until all
occurences of it are replaced. UI should warn the user that instances of such a
component must be updated, ideally providing ways to navigate to such locations,
and prevent execution until it is done.

On top of that, UI prepares geometry data of the whole user facing side of the
application that is to be passed to the renderer and initiates rendering of a
frame.

\subsection{Resolving Dependencies}
From the above we may conclude that the model is a critical component blocking
other modules except renderer and the design must be established early on to
further development. There is a blocking dependency of UI on the renderer as it
is impossible to display any results on-screen without a working renderer. This
makes the model and the renderer the highest priority goals. A simple bootstrap
renderer should be developed early on to unblock other modules. An improved
version would then be developed incrementally and concurrently with other
modules.

\section{Technical Evaluation}
The core technologies used in the project are:
\begin{itemlist}
  \item C++
  \item OpenGL
  \item GLSL
\end{itemlist}

Additionally, the following libraries are evaluated:
\begin{itemlist}
  \item anton{\_}core
  \item GLFW
  \item Dear Imgui
\end{itemlist}

\subsection{Evaluation of C++}
C++ is a widespread programming language with a large community of experts and a
standard committee. The language has been in use for over 30 years, especially
favoured by the game development industry for its performance, interoperability
as well as a range of frameworks and libraries. It is safe to assume that the
compilers of this language have been thoroughly tested for our use cases.

The project mandates use of a language that has a high interoperability with
graphics APIs. C++ fulfills this role is it allows for low-level manipulation of
memory, but at the same time provides facilities for high-level abstractions.
The complexity of the language is overwhelming, however, the commonly used
subset of the language is well known and should not pose any risk in terms of
possible target overruns. Using other less known languages, such as Rust, would
require an initial learning period taking away from the development and would
greatly contribute to stalls due to lack of essential knowledge furthering
overruns.

The version used is C++20 with the exception that only the C++17 feature set and
a subset of C++20 features is allowed. This is due to the fact that many of the
additions in C++20 have a major overhead in terms of compilation speed.

\subsection{Evaluation of OpenGL}
OpenGL (GL) is an industry standard graphics API in use for over 30 years. The
4.5 version of GL released in 2016 has been widely adopted with major GPU
retailers adding support for this version to their drivers within 2 years of the
release. Over 5 years of continuous support proves its stability and
reliability.

Although version 4.0 and especially 3.3 have a much longer tenure and larger
adoption, the newly added features in versions 4.1-4.5 are extremely convenient
and provide tools for the programmers, such as Debug Output (dbgout) which
outputs robust debug diagnostics on various errors easing development and
debugging.

GL provides two profiles: core and compatibility. The compatibility profile is
designed for upgrading existing software to newer versions of GL while
maintaining the existing codebase. We do not have any legacy code, hence
compatibility profile is not required and we may proceed with core.

An alternative to GL is Vulkan (Vk) which is a newer API targeted at ultrahigh
performance graphics programs. Primarily used in game development, Vk provides a
more robust API allowing a fine control over the hardware. Vk has several
disadvantages over GL for our purposes:
\begin{itemlist}
\item Vk is difficult. The concepts innate to Vk are difficult for beginners as
  it exposes the inner workings of the GPU drivers to a greater extent,
  primarily to allow for the fine control. An example is synchronisation which
  is done automatically in GL, but in Vk requires a considerable amount of work.
\item Vk is verbose. Compared to GL, the exact same boilerplate program requires
  several times more code to set up. Further development of the program in Vk
  would require proportionally more code hindering speed of development.
\end{itemlist}

Nebula is not a graphics intensive program, hence the benefits offered by Vk
are minimal while the disadvantages may significantly slow down the development
and cause target overruns.

Other mainstream options, such as Metal, DirectX, are not viable due to being
locked to specific platforms which are not out target platforms.

\subsection{Evaluation of GLSL}
GLSL, OpenGL Shading Language, is a shading language for creating GPU shading
programs (shaders) designed to work directly with GL. This compatibility
naturally translates to a simpler workflow as other options require
cross-compilation to intermediate formats to be able to load them. There are no
performance implications of choosing GLSL over other languages.

GLSL allows the users to choose a version of the compiler that will be used at
load time. To remain consistent with the version of GL, we use version '450
core'.

\subsection{Evaluation of anton{\_}core} TODO

\subsection{Evaluation of GLFW}
TODO

\subsection{Evaluation of Dear Imgui}
TODO

\end{document}
